\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-10-22 fully-abstract-comp/popl2019/paper.tex
Add a reference to Szabo's blogpost, small rephrasings.

In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although a design sketch for them is incorporated in the latest CHERI ISA reference~\citep{TODO}\todo{reference!} and related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997} and in technical documents.
% KJAA: Perhaps make it clear by who and when/where?
The hardware prevents these capabilities from being duplicated.
We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.

@@

Roughly speaking, this lemma says that under certain conditions, executing any executable capability under \srccm{} and \trgcm{} semantics will produce the same observable behavior.
The conditions require that the capability points to a memory region where code is loaded and that code must be either trusted and behave reasonably (i.e.\ respect the restrictions that \stktokens{} relies on, see Section~\ref{sec:well-form-reas}) or untrusted (in which case, it cannot have WBCF or LSE expectations, see \todo{explain}).

@@

Plenty of other papers enforce well-bracketed control flow at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
This includes particularly the line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.
This technique prevents certain classes of attacks by sanitizing addresses before direct and indirect jumps based on static information about a program's control graph and a shadow stack.
Contrary to \stktokens{}, CFI can be implemented on commodity hardware rather than capability machines.
However, its attacker model is different, and its security goals are weaker.
They assume an attacker that is unable to execute code but can overwrite arbitrary data at any time during execution (to model buffer overflows).
In terms of security goals, the technique does not enforce local stack encapsulation.
Also, it only enforces a weak form of control flow correctness saying that jumps stay within the static control flow graph of a program~\cite{Abadi2005Theory}.
Such a property ignores temporal properties and seems hard to use for reasoning.
There is also more and more evidence that these partial security properties are not enough to prevent realistic attacks in practice~\citep{Evans:2015:CJW:2810103.2813646,Carlini2015ControlFlowBending}.

>>>>

In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although a design sketch for them is incorporated in the latest CHERI ISA reference~\citep{TODO}\todo{reference!} and related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997,szabo_scarce_objects} and in technical documents.
% KJAA: Perhaps make it clear by who and when/where?
The hardware prevents these capabilities from being duplicated.
We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.

@@

Roughly speaking, this lemma says that under certain conditions, executing any executable capability under \srccm{} and \trgcm{} semantics will produce the same observable behavior.
The conditions require that the capability points to a memory region where code is loaded and that code must be either trusted and behave reasonably (i.e.\ respect the restrictions that \stktokens{} relies on, see Section~\ref{sec:well-form-reas}) or untrusted (in which case, it cannot have WBCF or LSE expectations, see Section~\ref{sec:well-form-reas}).

@@

Plenty of other papers enforce well-bracketed control flow at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
This includes particularly the line of work on control-flow integrity~\citep{abadi_control-flow_2005}.
This technique prevents certain classes of attacks by sanitizing addresses before direct and indirect jumps based on static information about a program's control graph and a shadow stack.
Contrary to \stktokens{}, CFI can be implemented on commodity hardware rather than capability machines.
However, its attacker model is different, and its security goals are weaker.
They assume an attacker that is unable to execute code but can overwrite arbitrary data at any time during execution (to model buffer overflows).
In terms of security goals, the technique does not enforce local stack encapsulation.
Also, it only enforces a weak form of control flow correctness saying that jumps stay within the static control flow graph of a program~\cite{Abadi2005Theory}.
Such a property ignores temporal properties and seems hard to use for reasoning.
There is also more and more evidence that these partial security properties are not enough to prevent realistic attacks in practice~\citep{Evans:2015:CJW:2810103.2813646,Carlini2015ControlFlowBending}.
